Dynamics of thin liquid film flows studied with the lattice Boltzmann method

### 2023-08-14 chapters/Paper_three.tex
Done for now

The wavelengths are obviously limited from above by the size of the computational domain. On the other extreme, too short wavelengths are prevented by the energetic cost of accomodating droplets on too small patches, since the local contact angle (that increases with decreasing $\lambda$~\cite{SuppMat}) would possibly exceed the largest substrate contact angle, such that the patterning itself would loose meaning. 

@@

Consequently, as we see from the inset of Fig.~\ref{fig:clusters_v0_sine}, where we plot the number of droplets $N(t)$ in time\footnote{A droplet is identified by the set (``cluster'') of points, in the plane, constituting each of the connected components of the set $\{\mathbf{x} \in [0,L]^2 | h(\mathbf{x},t) \geq h_{\ast}$\}; the clusters are determined by means of an algorithm of Hoshen-Kopelman type~\cite{HK}.}, in the steady state ($t \gg t_0$) $N(t)$ attains the value $N_{\infty} = 2(L/\lambda)^2$ (reported as horizontal lines), which equals the minima of Eq.~(\ref{eq:sinetheta}), for $v_{\theta}=0$.

@@

In order to better characterize the various morphologies we apply the theory of Minkowski's functionals. 
In particular, we employ the second order Minkowski structure metric, $q_2$~\cite{mickelShortcomingsBondOrientational2013, schallerPapaya22DIrreducible2020}, which can be computed from a Voronoi tessellation of the set of discrete points $(x_i, y_i)$ on the 2D lattice, such that the height field lies above a certain threshold\footnote{The expression is: $q_2 = \frac{1}{N}\sum_j \frac{1}{P_j}|\sum_k L^{(j)}_k e^{2i\phi^{(j)}_k}|$, where the inner sum runs over the edges of length $L^{(j)}_k$, of the $j$-th Voronoi cell, whose perimeter is $P_j$, and $\phi^{(j)}_k$ is the polar angle of the normal to the $k$-th edge. 
The outer sum represents an ensemble average over the $N$ points in the set.}. 
The $q_2$ metric quantifies the degree of anisotropy of the dewetting morphology, so it takes relatively large values if the structures formed display a preferential direction. 
Measuring $q_2$ then enables us to clearly distinguish between the formation of droplets and rivulets: much larger $q_2$ values are attained for the latter type of structure, as we can see in Fig.~\ref{fig:msm_q2}.
We observe, on the other hand, that such rivulets are metastable and eventually break up into droplets, as indicated by the collapse of $q_2$ at later times. 
Notice, though, that the $q_2$ signal for any $v_{\theta} >0$ always stays above the one for the static case, suggesting that even the smallest pattern velocity introduces a sizeable deformation of the spherical cap shape.
\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/Figure_5.pdf}
    \caption{MAIN PANEL: Rivulets life-times, $\tau_{\text{riv}}$, for various $\Gamma$ values.
    The dashed line is plotted as a guide to the eye to highlight the logarithmic dependence, in agreement with the theoretical prediction, Eq.~(\ref{eq:rivlt}).
    INSET: Height fluctuations, $\Delta h(t)$, vs time, computed along the rivulet axis, for three different $\Gamma$ values.
    }
    \label{fig:stab_ligs_lam2}
\end{figure}
The breakup is the result of a varicose mode of the rivulet~\cite{diezBreakupFluidRivulets2009, mechkovStabilityLiquidRidges2008}, whose wavelength is $\approx\lambda$, such that only $N_{\infty}/2$ droplets are counted after breakup. 
These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state"~\footnote{see movie file ligament\_formation\_and\_breakup.mp4 in~\cite{SuppMat}.}.

@@

\section{Supplemental Material}.
\subsection{Minkowski's structure metric \texorpdfstring{$q_2$}{hmm} in an extended parameter space.}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{graphics/SupMatFig_1.pdf}
    \includegraphics[width=0.4\textwidth]{graphics/SupMatFig_2.pdf}
    \caption{LEFT PANEL. Minkowski's structure metric $q_2$ for three different wavelengths $\lambda=64 h_0$, $\lambda=128 h_0$ and $\lambda=256 h_0$ (as in figure 4 of the main text) and for $\Gamma=1.5$ (all other parameters are as in the main text). 
    The time interval during which $q_2 \approx 1$ signals the emergence of rivulets, also for $\lambda = 64 h_0$ which is comparable with the spinodal wavelength $\lambda_s \approx 70 h_0$. 
    RIGHT PANEL. Comparison of the Minkowski's structure metric $q_2$ for $\delta\theta=5^{\circ}$ and $\delta\theta=10^{\circ}$, $\Gamma = 15$ and $\lambda = 256 h_0$.}
    \label{fig:q2_difflambda}
\end{figure}
\noindent In this section we test the robustness of the observation of the rivulet state over a wider parameter space and in particular at changing: 1) the 
pattern wavelength $\lambda$ and 2) the heterogeneity amplitude $\delta \theta$.
To this aim we have run simulations with $\Gamma=1.5$, $\lambda/h_0 = 64, 128$, and with $\Gamma = 15$, $\lambda = 256 h_0$, $\delta \theta = 5^{\circ}$, respectively.
In Fig.~\ref{fig:q2_difflambda} (left panel) we report the measurements of the Minkowski's structure metric $q_2$, whose increase from zero up to a plateauing value of $q_2 \approx 1$ indicates the emergence of rivulets, for the different $\lambda$'s ($\lambda=256 h_0$, which is the value considered in the main text, is also reported for comparison).
Analogously, in the right panel, we report $q_2$ as a function of time for $\delta \theta = 5^{\circ}$  and $\delta \theta = 10^{\circ}$ (the value used in the main text). 
We see that the $q_2$ metric attains the value $q_2 \approx 1$, signalling the rivulet state, albeit over a time interval shorter than for $\delta \theta = 10^{\circ}$, i.e. the rivulets lifetime decreases with $\delta \theta$. 
This was somehow expected, since obviously (and also in the static case) the patterning looses effectiveness as the contact angle mismatch is reduced (see, e.g.~\cite{konnurInstabilityMorphologyThin2000}).

>>>>

The wavelengths are obviously limited from above by the size of the computational domain. On the other extreme, too short wavelengths are prevented by the energetic cost of accomodating droplets on too small patches, since the local contact angle (that increases with decreasing $\lambda$~\ref{suppmat}) would possibly exceed the largest substrate contact angle, such that the patterning itself would loose meaning. 

@@

Consequently, as we see from the inset of Fig.~\ref{fig:clusters_v0_sine}, where we plot the number of droplets $N(t)$ in time\footnote{A droplet is identified by the set (``cluster'') of points, in the plane, constituting each of the connected components of the set $\{\mathbf{x} \in [0,L]^2 | h(\mathbf{x},t) \geq h_{\ast}$\}; the clusters are determined by means of an algorithm of Hoshen-Kopelman type~\cite{hoshenPercolationClusterDistribution1976}.}, in the steady state ($t \gg t_0$) $N(t)$ attains the value $N_{\infty} = 2(L/\lambda)^2$ (reported as horizontal lines), which equals the minima of Eq.~(\ref{eq:sinetheta}), for $v_{\theta}=0$.

@@

In order to better characterize the various morphologies we apply the theory of Minkowski's functionals. 
In particular, we employ the second order Minkowski structure metric, $q_2$~\cite{mickelShortcomingsBondOrientational2013, schallerPapaya22DIrreducible2020}, which can be computed from a Voronoi tessellation of the set of discrete points $(x_i, y_i)$ on the 2D lattice, such that the height field lies above a certain threshold\footnote{The expression is: $q_2 = \frac{1}{N}\sum_j \frac{1}{P_j}|\sum_k L^{(j)}_k e^{2i\phi^{(j)}_k}|$, where the inner sum runs over the edges of length $L^{(j)}_k$, of the $j$-th Voronoi cell, whose perimeter is $P_j$, and $\phi^{(j)}_k$ is the polar angle of the normal to the $k$-th edge. 
The outer sum represents an ensemble average over the $N$ points in the set.}. 
The $q_2$ metric quantifies the degree of anisotropy of the dewetting morphology, so it takes relatively large values if the structures formed display a preferential direction. 
Measuring $q_2$ then enables us to clearly distinguish between the formation of droplets and rivulets: much larger $q_2$ values are attained for the latter type of structure, as we can see in Fig.~\ref{fig:msm_q2}.
We observe, on the other hand, that such rivulets are metastable and eventually break up into droplets, as indicated by the collapse of $q_2$ at later times. 
Notice, though, that the $q_2$ signal for any $v_{\theta} >0$ always stays above the one for the static case, suggesting that even the smallest pattern velocity introduces a sizeable deformation of the spherical cap shape.
\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/Figure_5.pdf}
    \caption{MAIN PANEL: Rivulets life-times, $\tau_{\text{riv}}$, for various $\Gamma$ values.
    The dashed line is plotted as a guide to the eye to highlight the logarithmic dependence, in agreement with the theoretical prediction, Eq.~(\ref{eq:rivlt}).
    INSET: Height fluctuations, $\Delta h(t)$, vs time, computed along the rivulet axis, for three different $\Gamma$ values.
    }
    \label{fig:stab_ligs_lam2}
\end{figure}
The breakup is the result of a varicose mode of the rivulet~\cite{diezBreakupFluidRivulets2009, mechkovStabilityLiquidRidges2008}, whose wavelength is $\approx\lambda$, such that only $N_{\infty}/2$ droplets are counted after breakup. 
These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state"~\footnote{see movie file ligament\_formation\_and\_breakup.mp4 in~\ref{suppmat}.}.

@@

\section{Supplemental Material}.
\label{suppmat}
\subsection{Minkowski's structure metric \texorpdfstring{$q_2$}{hmm} in an extended parameter space.}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{graphics/SupMatFig_1.pdf}
    \includegraphics[width=0.4\textwidth]{graphics/SupMatFig_2.pdf}
    \caption{LEFT PANEL. Minkowski's structure metric $q_2$ for three different wavelengths $\lambda=64 h_0$, $\lambda=128 h_0$ and $\lambda=256 h_0$ (as in figure 4 of the main text) and for $\Gamma=1.5$ (all other parameters are as in the main text). 
    The time interval during which $q_2 \approx 1$ signals the emergence of rivulets, also for $\lambda = 64 h_0$ which is comparable with the spinodal wavelength $\lambda_s \approx 70 h_0$. 
    RIGHT PANEL. Comparison of the Minkowski's structure metric $q_2$ for $\delta\theta=5^{\circ}$ and $\delta\theta=10^{\circ}$, $\Gamma = 15$ and $\lambda = 256 h_0$.}
    \label{fig:q2_difflambda}
\end{figure}
\noindent In this section we test the robustness of the observation of the rivulet state over a wider parameter space and in particular at changing: 1) the 
pattern wavelength $\lambda$ and 2) the heterogeneity amplitude $\delta \theta$.
To this aim we have run simulations with $\Gamma=1.5$, $\lambda/h_0 = 64, 128$, and with $\Gamma = 15$, $\lambda = 256 h_0$, $\delta \theta = 5^{\circ}$, respectively.
In Fig.~\ref{fig:q2_difflambda} (left panel) we report the measurements of the Minkowski's structure metric $q_2$, whose increase from zero up to a plateauing value of $q_2 \approx 1$ indicates the emergence of rivulets, for the different $\lambda$'s ($\lambda=256 h_0$, which is the value considered in the main text, is also reported for comparison).
Analogously, in the right panel, we report $q_2$ as a function of time for $\delta \theta = 5^{\circ}$  and $\delta \theta = 10^{\circ}$ (the value used in the main text). 
We see that the $q_2$ metric attains the value $q_2 \approx 1$, signalling the rivulet state, albeit over a time interval shorter than for $\delta \theta = 10^{\circ}$, i.e. the rivulets lifetime decreases with $\delta \theta$. 
This was somehow expected, since obviously (and also in the static case) the patterning looses effectiveness as the contact angle mismatch is reduced (see, e.g.~\cite{konnurInstabilityMorphologyThin2000}).

### 2023-08-14 chapters/Paper_two.tex
Done for now

More in general, thermal fluctuations can accelerate the appearance of film instabilities~\cite{Rauscher2008,tsekov1993effect, fetzerThermalNoiseInfluences2007, zhangMolecularSimulationThin2019}.

>>>>

More in general, thermal fluctuations can accelerate the appearance of film instabilities~\cite{rauscherWettingPhenomenaNanofluidics2008,tsekovEffectThermalFluctuations1993, fetzerThermalNoiseInfluences2007, zhangMolecularSimulationThin2019}.
